### 2013-08-31 msc.tex
initial

\documentclass[a4paper,11pt]{article}
%\usepackage{algorithmicx}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{placeins}
\usepackage{amsthm}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usepackage{comment}
\usetikzlibrary{arrows,decorations.pathreplacing}
\newcommand{\atomic}[4]{\node(lab#1) at (#3,-#4+0.5) {#2};\draw (#3,-#4-0.1) to (lab#1);\node[inner sep=0pt](#1) at (#3,-#4) {};}
\algblockdefx{Struct}{EndStruct}[1]{\textbf{struct} \textsc{#1}}{}
\algnewcommand{\algorithmicgoto}{\textbf{go to}}
\algnewcommand{\Goto}{\textbf{goto} }
\algnewcommand{\Label}{\State\unskip}
\def\none{\textbf{none} }
\def\T{\ensuremath{\operatorname{\mathcal{T}}}\text{ }}
\def\int{\ensuremath{\operatorname{\textbf{int}}}}
\algnotext{EndStruct}
\newtheorem{lemma}{Lemma}
\newtheorem{observation}{Observation}
\newtheorem{theorem}{Theorem}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\fn}[1]{\textsc{#1}}
\usepackage{float}
\newfloat{Listing}{t}{lop}
\author{Robert Obryk}
\title{Write-and-f-array: implementation and an application}
\date{}
\begin{document}

\maketitle
\begin{abstract}
	We introduce a new shared memory object: the write-and-f-array, provide its wait-free implementation and use it to construct an improved wait-free implementation of the fetch-and-add object. The write-and-f-array generalizes single-writer write-and-snapshot\cite{write-and-snap} object in a similar way that the f-array\cite{f-array} generalizes the multi-writer snapshot object. More specifically, a write-and-f-array is parametrized by an associative operator $f$ and is conceptually an array with two atomic operations:

\begin{itemize}
\item write-and-f modifies a single array's element and returns the result of applying $f$ to all the elements,
\item read returns the result of applying $f$ to all the array's elements.
\end{itemize}

We provide a wait-free implementation of an $N$-element write-and-f-array with $O(N \log N)$ memory complexity, $O(\log^3 N)$ step complexity of the write-and-f operation and $O(1)$ step complexity of the read operation. The implementation uses CAS objects and requires their size to be $\Omega(\log M)$, where $M$ is the total number of write-and-f operations executed. We also show, how it can be modified to achieve $O(\log^2 N)$ step complexity of write-and-f, while increasing the memory complexity to $O(N \log^2 N)$.

The write-and-f-array can be applied to create a fetch-and-add object for $P$ processes with $O(P \log P)$ memory complexity and $O(\log^3 P)$ step complexity of the fetch-and-add operation. This is the first implementation of fetch-and-add with polylogarithmic step complexity and subquadratic memory complexity that can be implemented without CAS or LL/SC objects of unrealistic size\cite{ellen-fai}.
\end{abstract}
\clearpage
\section{Introduction}

\subsection{Concurrent objects}
A useful abstraction in a single-threaded system is an object: an entity with a set of methods one can invoke
and a specification of semantics of these methods. Naturally, a program can use multiple objects and invoke
their methods in any order.

We want to use a similar abstraction to model interaction between threads in a multi-threaded system.
Consider a set of independent, concurrently running threads (running possibly
distinct code) that can only communicate by calling methods on some objects. We place no assumptions on
the speed of execution and delays of threads -- they may wait arbitrarily long before a method call.
We also assume that method calls on the objects complete instaneously and that no two of them happen at the same time.
This allows us to define semantics of the objects in the same fashion we define them in the single-threaded case:
method of every object are invoked in a known order, so single-threaded semantics suffice to determine object's behaviour\footnote{The whole program may still be nondeterministic, due to scheduling vagaries.}.
We will call such objects concurrent objects to emphasize that they may be accessed concurrently.

%An asynchronous shared memory system consists of a number of threads and a number of shared objects. Each
%thread is a Turing-complete automaton that can addiationally execute external actions on the shared objects.
%We must provide the semantics of such a system: we will do this by defining a process that yields all possible
%correct executions of a program in this system (the system is nondeterministic).

%A correct execution trace is a sequence of external actions of all threads. When we take a subsequence that
%corresponds to a single object, the return values of the actions must satisfy object's semantics. When we
%take a subsequence that corresponds to a single thread, an execution of that thread alone that would receive
%these return values from the actions must initiate a sequence of actions that starts with exactly these actions.
%Time will refer from now on to executio traces?

%Informally speaking, threads can interleave their external actions arbitrarily: they can ,,hang'' for long
%periods of time, or execute many actions without interruption; they may even be ,,hung'' forever (we will call
%such threads stopped).

An object we will predominantly use in this paper is a CAS\footnote{Compare-And-Swap} register.
We will specify its semantics by providing pseudocode which correctly implements this object in
a single-threaded program (this will be our method of choice of providing object's semantics):

\begin{algorithmic}[1]
\State \T $x$

\Function{rd}{}
	\State\Return $x$
\EndFunction

\Function{wr}{$y$}
	\State $x \gets y$
\EndFunction

\Function{CAS}{$o, n$}
\If{$x = o$}
	\State $x \gets n$
	\State\Return true
\Else
	\State\Return false
\EndIf
\EndFunction
\end{algorithmic}

Intuitively, such a register holds a value that can be read (by \fn{rd}), modified unconditionally (by \fn{wr})
and modified conditionally (by \fn{CAS}). It's interesting due to its universality properties and because it's
commonly seen in real-world hardware. % T?

\begin{example}
Let's consider two threads, executing following pseudocodes (X and Y are two CAS objects, with initial value 0):

\begin{enumerate}
\item\begin{algorithmic}[1]
\State X.\fn{wr}(1)
\State Y.\fn{rd}()
\end{algorithmic}

\item\begin{algorithmic}[1]
\State Y.\fn{wr}(1)
\State X.\fn{rd}()
\end{algorithmic}
\end{enumerate}

Figure~xx presents a possible execution of such two threads. Note that in any correct execution,
at least one of the \fn{rd}() calls returns 1.
\end{example}

\subsection{Serialization}
In many situations it would be convenient to have more complicated concurrent objects: counters that can be incremented,
queues that can be pushed to and popped from, etc. Alas, real-world hardware usually doesn't implement such objects
directly. We will thus implement them using CAS registers or other simpler objects: we will substitute a set of simpler objects for every such object and for every method call on this objects, execute a procedure that implements it instead. Obviously, we will want such substitution to preserve semantics of the program in which it was substituted. The following condition immediately implies this
\footnote{For a detailed discussion of subtleties related to this definition, refer to chapter 3.3-3.6 of TheArt}:

\begin{definition}
An implementation of a concurrent object \emph{serializes}, if for every execution of a program with the implementation
substituted for the object:

For every call to object's method that completes, we can define a \emph{serialization point} in its execution
interval, such that all these calls would return the same values if they happened instaneously at their serialization
points.
\end{definition}

An observant reader might note that if all procedures in an implementation never terminate, the definition is
vacously true. Indeed, we will present some terminations guarantees separately, but first let us consider an
example of an implementation of a simple structure.

%This paper's contribution is an implementation of a novel concurrent object using CAS registers.

%\begin{definition}
%An implementation of an object of type X using objects of type Y is a 
%That is, we can provide procedures that can be executed instead of invoking external actions of such a counter or
%a queue; these procedures use a set of shared CAS registers and, informally, can be substituted for these actions
%without a change in behaviour. More formally:

%An implementation of a shared object serializes

% nonblocking

\begin{example}
Let's consider an increase-only counter: a shared object with the following semantics:

\begin{algorithmic}[1]
\State \textbf{int} $x$

\Function{inc}{}
	$x \gets x + 1$
\EndFunction
\Function{read}{}
	\State\Return$x$
\EndFunction
\end{algorithmic}

We will see that the following implementation serializes to this shared object:

\begin{algorithmic}[1]
\State \textbf{int} $V$ \Comment{A CAS object holding an \textbf{int}}

\Function{inc}{}
	\Repeat
		$o \gets V.\fn{rd}()$
		\Until{$V.\fn{CAS}(o, o+1)$}
\EndFunction

\Function{read}{}
	\State\Return $V.\fn{rd}()$
\EndFunction
\end{algorithmic}

Note that every call to \fn{inc} executes exactly one successful call to V.CAS. Let us claim that the serialization
point of \fn{inc} is that successful call to CAS and the serialization point of \fn{read} is the call to
V.READ. We can easily see that the number returned by READ is exactly the number of \fn{inc}s that serialized
before that \fn{read}.
% figure with serialization
\end{example}

We can also see that this implementation satisfies some global progress property -- when the CAS in \fn{inc} fails,
another \fn{inc} must have suceeded very recently. In fact, it must have succeeded between the most recent execution
of xx and the failed CAS. Still, a single \fn{inc} operation can fail to terminate.

\subsection{Wait-freeness}

Consider an alternative implementation of an increment-only counter:

\begin{algorithmic}[1]
\State \textbf{int} $T[P]$ \Comment{T is an array of $P$ CAS objects of type int}
\Function{inc$_p$}{}
	\State $x \gets T[p].\fn{rd}()$
	\State $T[p].\fn{wr}(x + 1)$
\EndFunction
\Function{read}{}
	\State $R \gets 0$
	\For{$i \gets 1 \mbox{to} P$}
		\State $R \gets R + T[i]$
	\EndFor
	\State\Return $R$
\EndFunction
\end{algorithmic}

Here we differentiate between threads: the index p in \fn{inc} corresponds to the current thread's ID from range
$\{0, \ldots, P-1\}$. Let's see first that this implementation actually serializes to the increment-only counter.
Obviously, we need to choose the call to WRITE as the serialization point of \fn{inc}. From the monotonicity
of entries of T we can infer that \fn{read} returns a value between (inclusive) the number of calls to \fn{inc} that have
serialized before \fn{read} has started and finished. Thus if a call to \fn{read} returns x, there is a point during its
execution when exactly x \fn{inc}s had been serialized. We can choose this point to be the serialization point of
\fn{read}\footnote{Note that it need not correspond to an action executed by the implementation of \fn{read}}.

We can see that this implementation satisfies a very strong termination condition -- the number of steps every
procedure takes is bounded and the bound depends only on the number of threads (and not on the properties of
a particular execution trace). We will call such implementations \emph{wait-free}. For a discussion of weaker
termination guarantees, see chapter 3.7 of TheArt.

Every object that has a single-threaded specification has a wait-free implementation\ref{universal-wf} using
only CAS registers.
However, time complexities of the operations are larger by at least a factor of number of threads. 
Thus, creating faster wait-free implementations of data structures is interesting, also from a practical
point of view.

\subsection{Snapshots}

The object implemented in this paper is a generalization of a popular building block -- an atomic snapshot
object\cite{snapshot-idea}. Its semantics can be defined by following single-threaded pseudocode:

\begin{algorithmic}[1]
\State \T $T[N]$
\Function{update$_i$}{$x$}
	\State $T[i] \gets x$
\EndFunction
\Function{scan}{}
	\For{$i \gets 0 \mbox{to} N$}
		\State $R[i] \gets T[i]$
	\EndFor
	\State\Return $R$
\EndFunction
\end{algorithmic}

Intuitively, this object represents an array that can be modified piecewise and read all at once.
There are known wait-free implementations of an $N$-element atomic snapshot with $O(1)$ \fn{update} and $O(N)$ \fn{scan} step complexities\cite{snapshot-impl}. These complexities are obviously
optimal.

One can try to improve on this structure in two different ways. First, we can merge the update and scan operations
into an operation that does an update and immediately thereafter a scan. This object, which we'll call write-and-snapshot\cite{write-and-snap}, can be defined by the following single-threaded specification:

\begin{algorithmic}[1]
\State \T $T[N]$
\Function{update-and-snap$_i$}{$x$}
	\State $T[i] \gets x$
	\For{$i \gets 0 \mbox{to} N$}
		\State $R[i] \gets T[i]$
	\EndFor
	\State\Return $R$
\EndFunction
\end{algorithmic}

For the second improvement, let us note that in many cases, one doesn't need the \fn{scan} operation to return the whole array, but rather some sort of its digest. The f-arrays\cite{f-array} allow one to do precisely that for digests that can be expressed as
a result of applying an associative operation to all the array's elements (for example, if we need to retrieve the sum of all array's elements). If we call the operator $f$, the following is a specification of an f-array:

\begin{algorithmic}[1]
\State \T $T[N]$
\Function{update$_i$}{$x$}
	\State $T[i] \gets x$
\EndFunction
\Function{scan}{}
	\State\Return $f(T[0], T[1], \ldots, T[N-1])$
\EndFunction
\end{algorithmic}

There is an implementation of f-array with step complexities of \fn{update} and \fn{scan} being respectively $O(\log N)$ and $O(1)$ and the memory complexity
being linear in $N$.

This paper introduces a write-and-f-array: an object that combines these two modifications. Its semantics are defined by the single-threaded specification in Listing~\ref{spec-general}. We then provide implementation of a single-writer
write-and-f-array: that is, one that disallows concurrent modifications to the same array element.

\begin{Listing}
\begin{algorithmic}[1]
\State \textbf{int} V
\Function{fetch-and-add}{$x$}
	\State $r \gets V$
	\State $V \gets V + x$
	\State\Return $r$
\EndFunction
\end{algorithmic}
\caption{Specification of semantics of fetch-and-add}
\label{faa-spec}
\end{Listing}

We then use this object to implement a fetch-and-add object. Fetch-and-add is a generalization of the increment-only counter we've seen. Its semantics are specified in Listing~\ref{faa-spec}.
Fetch-and-add objects can be used to produce unique identifiers, implement
mutual exclusion, barrier synchronization\cite{faa-sync}, or work queues\cite{faa-queue}. Its known wait-free shared memory implementations for $P$ processes have $\Omega(P^2)$ memory complexity and $O(\log^2 P)$ step complexity
of the operation \cite{ellen-fai}\cite{closed-object}. They also need to employ complicated memory management techniques to bound their memory use. Our implementation reduces the memory complexity to $O(P\log P)$ while maintaining polylogarithmic step complexity.

%In the second section of this paper we construct a helper structure, while in the third section we construct an implementation of the write-and-f-array. In the last section we describe possible extensions of this
%structure and open problems.

\section{The History Object}

\begin{Listing}[!h]
\begin{algorithmic}[1]
	\State $\T H[]$
	\State $\int ver$
	\Function{get-current}{}
		\State \Return $ver$, $H[ver]$
	\EndFunction
	\Function{get}{$v$}
		\If{$v \leq ver - N$}
			\State \Return \none
		\Else
		\State \Return $H[ver]$
		\EndIf
	\EndFunction
	\Function{publish}{$v$, $T$}
		\If{$v = ver + 1$}
		\State $H[v] \gets T$
			\State $ver \gets v$
			\State \Return true
		\Else
			\State \Return false
		\EndIf
	\EndFunction
\end{algorithmic}
\caption{Specification of the history object}
\label{hist-spec}
\end{Listing}
We will first implement a helper object -- the history object. Intuitively, it contains a versioned memory cell and allows retrieval of past $N$ values of the cell. The cell holds objects of type \T.
Our wait-free implementation will impose an additional constraint on its use: every call to \fn{publish} is parametrized by an integer in range $[0,P)$ and executions of \fn{publish$_p$} for the same $p$ can not run simultaneously.
	Both $N$ and $P$ are known when the object is created. The exact semantics of the structure are given by the single-threaded specification in Listing~\ref{hist-spec}

% refer to impl
\FloatBarrier

Our implementation is presented below:

\begin{algorithmic}[1]
	\State $\textbf{pair}\left<\int v, \int p\right> S$ \Comment{Version number and $p$ of the most recently published value}
	\State $\textbf{pair}\left<\int v, \T T\right> H[N]$ \Comment{A circular buffer of last $N$ values (except maybe the most recent one)}
	\State $\textbf{pair}\left<\int v, \T T\right> L[P]$ \Comment{Temporary storage for values being published}
	\Function{get-current}{}
		\State $s \gets S$ \Comment{Possible serialization point}
		\State \fn{help}()
		\State \Return $H[s.v \mod N]$
	\EndFunction
	\Function{help}{}
		\State $s \gets S$
		\State $l \gets L[s.p]$
		\State $h \gets H[s.v \mod N]$
		\If{$l.v = s.v \textbf{and} h.v < s.v$}
		\State $\text{CAS}(H[s.v \mod N], h, l)$ \label{hist-help-cas}
		\EndIf
	\EndFunction
	\Function{get}{v}
		\State $s \gets S$ \label{hist-get-notyet-sp} \Comment{Possible serialization point}
		\If{$s.v < v$}
		\State \Return \none \label{hist-get-notyet} \Comment{The requested version hasn't been published yet}
		\EndIf
		\State \fn{help}()
		\State $h \gets H[v \mod N]$
		\If{$h.v = v$}
			\State \Return $h.T$
		\Else
			\State \Return \none \label{hist-get-tooold}
		\EndIf
	\EndFunction
	\Function{publish$_p$}{v, T}
		\State $s \gets S$ \label{hist-pub-earlier-sp} \Comment{Possible serialization point}
		\If{$v \neq s.v + 1$}
			\State \Return \textbf{false} \label{hist-pub-earlier-exit}
		\EndIf
		\State \fn{help}()
		\State $L[p] \gets (v, T)$ \label{hist-pub-latest}
		\If{$\text{CAS}(S, s, (v, p)$) \textbf{failed}} \Comment{Possible serialization point}\label{hist-pub-later-sp}
			\State \Return \textbf{false}
		\EndIf
		\State \Return \textbf{true}
	\EndFunction
\end{algorithmic}

As the comments indicate, $L[p]$ is used to temporarily hold the value being published by \fn{publish$_p$}. The array $H$ is used to hold, roughly,
the $N$ most recently published values (the most recently published value might be absent; exact semantics will be given by the lemmas below) together with their version numbers. The $S$ field holds the particulars of the most recently published value; the successful publish calls will serialize at the moment S changes. Our implementation contains an additional function \fn{help}. It is used internally to write the most recently published value to the $H$ array, if it isn't stored there already.

We will now prove some properties related to exact semantics of $H$ and $L$ and then use them to prove that our implementation serializes to the specification.

%\begin{Listing}
%\caption{History data structure specification}
%\label{hist-spec}
%\end{Listing}

We posit that the serialization point of \fn{publish} is in~line~\ref{hist-pub-later-sp}, unless that line isn't reached. In that case (ie. when the call returns false in~line~\ref{hist-pub-earlier-exit})
the serialization point is in~line~\ref{hist-pub-earlier-sp}.

%We posit the following serialization points: get\_ver always has its serialization point in line~\ref{hist-getver-sp}. get's serialization point is in line~\ref{hist-get-sp} unless it has returned None
%in line~\ref{hist-get-notyet}; its serialization point is then in line~\ref{hist-get-notyet-sp}. Likewise, publish's serialization point is in line~\ref{hist-pub-later-sp} unless it has returned false
%in~line~\ref{hist-pub-earlier-exit}; then it is in line~\ref{hist-pub-earlier-sp}.

\begin{lemma}\label{hist-cv-latest}
If $S = (v, p)$ at time $t$ and at some later point in time $L[p] = (v, T)$, then there was a successful call to \fn{publish$_p$}$(v, T)$ with serialization point before time $t$.
\end{lemma}
\begin{proof}
S must have been modified by a successful CAS in line~\ref{hist-pub-later-sp}. Obviously, a successful invocation of \fn{publish$_p$}$(v, T)$ must have executed that CAS. Let's denote it by $\pi$.
This invocation sets $L[p]$ to $(v, T)$. What remains to be proven is that no later invocation of \fn{publish$_\cdot$} will set $L[p]$ to $(v, x)$ for any x.
Only \fn{publish$_p$} invocations modify $L[p]$ and they must start only after $\pi$ has finished. Together with a simple observation that $S.v$ is nondecreasing in time this implies
that any later invocation of \fn{publish} that sets $L[p]$ must have been called with a strictly greater $v$.
\end{proof}

\begin{observation}
\label{obs-hist-is-correct}
If $H[i] = (v, T)$ at some point then a successful call to \fn{publish$_\cdot$}$(v, T)$ has had its serialization point earlier.
\end{observation}

\begin{lemma}\label{lem-hist-help-comes}
Let $\pi$ be a call to \fn{publish$_p$}. Let $\pi_q$ be a successful call to \fn{publish$_q$} and $\pi_p$ a call to \fn{publish$_p$}, both with serialization points after serialization point of $\pi$. Then there is a call to \fn{help} that starts after serialization point of $\pi$ and ends before both serialization point of $\pi_q$ and execution of line~\ref{hist-pub-latest} in $\pi_p$.
\end{lemma}
\begin{proof}
The serialization point of $\pi$ must happen before $\pi_q$ executes line~\ref{hist-pub-earlier-sp} -- if S changed between line~\ref{hist-pub-earlier-sp} and line~\ref{hist-pub-later-sp} of $\pi_q$, $\pi_q$ would fail.
Thus the call to \fn{help} from $\pi_q$ will start after serialization point of $\pi$ and will finish before the serialization point of $\pi_q$. The call to \fn{help}
from $\pi_p$ will start after serialization point of $\pi$ ($\pi_p$ may only start after $\pi$ has finished) and will finish before line~\ref{hist-pub-latest} of $\pi_p$.
The call to \fn{help} that finishes earlier out of these two satisfies both conditions.
\end{proof}

\begin{lemma}
\label{lem-hist-is-complete}
If a call to \fn{help}() has executed fully (ie. started and finished) after the serialization point of a successful \fn{publish$_\cdot$}$(v, \cdot)$, then $H[v\mod N].v \geq v$.
\end{lemma}
\begin{proof}
Let us first note that $H[v\mod N].v$ is nondecreasing.
It thus suffices to prove that the condition is met at some point before the end of the call to \fn{help}. We will do so by induction on serialization point of the \fn{publish$_\cdot$}.
Consider a call to \fn{publish$_p$}$(v, \cdot)$ that serializes at time $t$ and a call to \fn{help} that starts after $t$. By Lemma~\ref{lem-hist-help-comes} there is a call to \fn{help} that finishes before next serialization point of a successful \fn{publish$_\cdot$} (thus before $S$
is modified) and before $L[p]$ is modified. Without loss of generality we can assume that the call to \fn{help} we are considering satisfies these conditions. By inductive hypothesis and Lemma~\ref{lem-hist-help-comes}, since a successful call to \fn{publish$_\cdot$}$(v-N+1, \cdot)$ has occured before time $t$, $H[v \mod N].first \geq v-N$ at time $t$. Taking into account that $H[i].v \mod N = i$, $H[v \mod N].v = v-N$ or $H[v \mod N] \geq v$ at any later point in time.
Thus CAS in line~\ref{hist-help-cas} either succeeds, which causes the lemma's conclusion to start being satisfied, or fails, which means that the conclusion was already satisfied.
\end{proof}

\begin{theorem}
The operations in the implementation of the history object serialize to the single-process structure with serialization points of \fn{publish$_\cdot$} as posited earlier.
\end{theorem}
\begin{proof}
Note that $S.v$ is exactly the first argument of the latest successful \fn{publish$_\cdot$}. Let us first consider a call to \fn{get-current} that returns $r$. From Lemma~\ref{lem-hist-is-complete} we know that 
$r.v \geq v.v$.
By Observation~\ref{obs-hist-is-correct}, \fn{publish$_\cdot$}$(r.v, \cdot)$ happened before the read from $H[v.v\mod N]$. If $r.v = v.v$, then \fn{publish$_\cdot$}$(r.v, \cdot)$ was the most recent successful \fn{publish$_\cdot$} at the time when $S$ was read, so we can serialize the operation there. Otherwise, \fn{publish$_i$}$(r.v, \cdot)$ has happened after that read, so we can serialize the operation just after it.

Let us consider a calls to \fn{publish$_\cdot$}:
\begin{itemize}
\item If a call to \fn{publish$_\cdot$}$(v, \cdot)$ fails in line~\ref{hist-pub-earlier-exit}, then we can see that the most recently serialized (from the point of view of line~\ref{hist-pub-earlier-sp}) successful \fn{publish$_\cdot$} published a version different than $v-1$.
\item If a call to \fn{publish$_\cdot$}$(v, \cdot)$ fails by failing the CAS in line~\ref{hist-pub-later-sp}, then a successful \fn{publish$_\cdot$} has occured after line~\ref{hist-pub-earlier-sp} had been executed, so the most recent successful \fn{publish$_\cdot$} at the time of CAS has published a version later than $v-1$.
\item If a call to \fn{publish$_\cdot$}$(v, \cdot)$ succeeds, then at the time of CAS in line~\ref{hist-pub-later-sp} the most recently published version is $v-1$.
	\end{itemize}
	This suffices to prove that the value returned by \fn{publish$_\cdot$} is always correct wrt. posited
	serialization order. What remains to consider are the calls to \fn{get}:
\begin{itemize}
	\item If a call to \fn{get}$(v)$ fails in line~\ref{hist-get-notyet}, then the most recent successful \fn{publish$_\cdot$} when line~\ref{hist-get-notyet-sp} executed had a smaller version number than requested, so we can serialize the call to \fn{get} at this line.
	\item If a call to \fn{get}$(v)$ fails in line~\ref{hist-get-tooold} because $h.v > v$, then (by Observation~\ref{obs-hist-is-correct}) a successful call to \fn{publish$_\cdot$}$(h.v, \cdot)$ has occured by that time and $h.v \geq v+N$, so if we serialize \fn{get} at that point, it should fail.
	\item If a call to \fn{get}$(v)$ fails in line~\ref{hist-get-tooold} because $h.v < v$, then (by Lemma~\ref{lem-hist-is-complete}) a call to \fn{publish$_\cdot$}$(v, \cdot)$ has not occured before the call to \fn{help} in line~X. We can thus serialize this \fn{get} just before the call to \fn{help} has started.
	\item If a call to \fn{get}$(v)$ succeeds, then by Observation~\ref{obs-hist-is-correct} it returns a value that was published by a successful call to \fn{publish$_\cdot$}$(v, \cdot)$. By Lemma~\ref{lem-hist-is-complete}, version $v+N$ was not published before the call to \fn{help} has started. Thus we can serialize this \fn{get} just after the successfull call to \fn{publish$_\cdot$}$(v, \cdot)$ has occured, or, if it has occured before \fn{get} started, just before the call to \fn{help} started.
\end{itemize}
\end{proof}

Obviously all operations run in $O(1)$ time. Memory complexity is $O(N+P)$, where $N$ and $P$ are the parameters defined at the beginning of this section.
We will now use this object in our main result, the implementation of a \textsc{write-and-f-array}.
\section{The write-and-f-array}

We will now present our main result -- an implementation of write-and-f-array.
We will present a wait-free implementation of a slightly richer structure. The additional operations and return values are required due to its recursive construction.
A single-threaded specification of the structure is presented in Listing~\ref{spec}. The \fn{version} is a black-box integer-valued function, subject to following conditions:
\begin{enumerate}
	\item Subsequent return values of \fn{version} are nondecreasing.
	\item If a call to \fn{version}(\textbf{false}) is followed by a call to \fn{version}(\textbf{true}), the second call must return a strictly greater integer than the first one.
\end{enumerate}
The second condition causes version numbers to group calls to \fn{write-and-f} into groups, such that no \fn{read} returns a value with only some updates from a group applied.
%reason for version
The concurrent implementation will be restricted by disallowing concurrent calls to \fn{write-and-f$_i$} for the same $i$.

\begin{Listing}
\begin{algorithmic}[1]
	\State \T $v[N]$
	\State \textbf{int} $\text{last\_update}[N]$
	\State \textbf{int} $\text{last\_version}[N]$
	\State \textbf{T} $\text{last\_val}[N]$

	\Function{update$_i$}{$T$}
		\State $r \gets f(v[0], v[1],\ldots, v[N-1])$
		\State $v[i] \gets T$
		\State $\text{last\_update}[i] \gets \text{last\_update}[i] + 1$
		\State $\text{last\_version}[i] \gets \fn{version}(\textbf{true})$
		\State $\text{last\_val}[i] \gets r$
		\State \Return $\text{last\_update}[i], \text{last\_version}[i], \text{last\_val}[i]$
	\EndFunction
	\Function{get-last}{$i$}
		\State \Return $\text{last\_update}[i], \text{last\_version}[i], \text{last\_val}[i]$
	\EndFunction
	\Function{read}{}
		\State \Return $\fn{version}(\textbf{false}), f(v[0], v[1],\ldots, v[N-1])$
	\EndFunction
\end{algorithmic}
\caption{The enriched write-and-f-array specification}
\label{spec}
\end{Listing}

Our implementation is conceptually very similar to Jayanti's tree-based f-array implementation. It uses a binary tree structure, with each array element assigned to a leaf.
We will construct it recursively. The implementation for $N = 1$ is presented in Listing~\ref{impl1}. If we note that no concurrent calls to \fn{write-and-f} may be made in it, we can easily
see that it is indeed correct and that all operations take constant time.

\begin{Listing}[!h]
\begin{algorithmic}[1]
	\State \textbf{pair<int v, \T T>} S
	\Function{write-and-f$_0$}{T'}
		\State v, T $\gets$ S
		\State S $\gets$ (v + 1, T')
		\State \Return v + 1, v + 1, T
	\EndFunction
	\Function{get-last}{$i$}
		\State v, T $\gets$ S
		\State \Return v, v, T
	\EndFunction
	\Function{read}{}
		\State \Return S
	\EndFunction
\end{algorithmic}
\caption{The enriched write-and-f-array snapshot implementation for $N=1$}
\label{impl1}
\end{Listing}
\FloatBarrier

The implementation for $N > 1$ is presented below (interspersed with comments):
\begin{algorithmic}[1]
	\Struct{node-value}
		\State \T $T[2]$ \Comment{Child values}
		\State \textbf{int} $v[2]$ \Comment{Child versions}
	\EndStruct
	\Struct{last-value}
		\State \textbf{int} $n$
		\State \textbf{int} $v$
		\State \T $T$
	\EndStruct
	\State \textsc{write-and-f-array} $C[2]$ \Comment{of size resp. $\lceil{}N/2\rceil{}$ and $\lfloor{}N/2\rfloor{}$}
	\State \textbf{history object<node-value>} $H$ \Comment{of size $N+1$, for $N$ concurrent updates}
	\State \textbf{last-value} $L[N]$ %\Comment{thread's update count, version, value from before update}
	\algstore{impl}
\end{algorithmic}

C[0] and C[1] are the subobjects from the recursive construction -- they are of size $\lceil{}N/2\rceil{}$ and $\lfloor{}N/2\rfloor{}$, respectively. Array elements of the enclosing
structure are mapped bijectively to the array elements of these two subobjects. Let's denote the mapping by $side(i)$ and $child\_id(i)$: the element $i$ is mapped to element $child\_id(i)$ in C[$side(i)$].
History object H is used to store the object's current value (the value that would be returned by \fn{read}) and past values. The version numbers exported by the history object correspond to the values returned by \fn{version}
in the specification. The elements of H aren't just values; they contain instead values of both children together with their versions.
The array L is used to store the values \fn{get-last} should return, but the values there might be stale (we will prove bounds on their staleness later on).

\begin{algorithmic}[1]
	\algrestore{impl}
	\Function{read}{}
		\State $v, h \gets H.\fn{get-current}()$ \label{get-getownver}
		\State \Return $v, f(h.T[0], h.T[1])$
	\EndFunction
	\Function{update$_i$}{}
		\State $v, h \gets H.\fn{get-current}()$ \label{upd-getown}
		\State $h'.v[0], h'.T[0] \gets C[0].\fn{read}()$ \label{upd-getch-0}
		\State $h'.v[1], h'.T[1] \gets C[1].\fn{read}()$ \label{upd-getch-1}
		\State $\fn{help}(v \mod N)$
		\State \Return $H.\fn{publish$_i$}(v+1, h')$ \label{upd-publish}
	\EndFunction
	\Function{write-and-f$_i$}{T}
		\State $C[side(i)].\fn{write-and-f$_{child\_id(i)}$}(T)$ \label{upd-chupd}
		\If{\textbf{not} \fn{update$_i$}()} \label{updcall1}
		\State \fn{update$_i$}() \label{updcall2}
		\EndIf
		\State \Return $\fn{get-last}(i)$
	\EndFunction
	\algstore{impl}
\end{algorithmic}

The implementation of \fn{read} and \fn{write-and-f} strongly resemble the f-array. \fn{write-and-f} uses a helper function \fn{update}, that intuitively pushes new values from $C[0]$ and $C[1]$ to $H$.
We will show that it suffices to attempt to call \fn{update} twice to accomplish that.

\begin{algorithmic}[1]
	\algrestore{impl}
	\Function{help}{x}
		\State $l_c \gets C[side(x)].\fn{get-last}(child\_id(x))$ \label{get-last-recurse}
		\State $v_{current} \gets H.\fn{get-current}().v$
		\State binary search $\{v_{current}-(N+1),\ldots, v_{current}\}$ for first such $v$ that: \label{help-binsch}
		\Statex $H.\fn{get}(v) \neq \none \textbf{ and } H.\fn{get}(v).v[side(x)] \geq l_c.v$ %indent!
		\If{no such $v$ exists}
			\State \Return
		\EndIf
		\State $h_{old} \gets H.\fn{get}(v-1)$ \label{help-getprev}
		\State $h_{new} \gets H.\fn{get}(v)$
		\If{$h_{old} = \none \textbf{ or } h_{new} = \none$}
			\State \Return
		\EndIf
		\If{$side(x) = 0$}
		\State $T \gets f(l_c.T, h_{old}.T[1])$
		\Else
		\State $T \gets f(h_{new}.T[0], l_c.T)$
		\EndIf
		\State $l \gets L[x]$
		\If{$l.n < l_C.n$} \label{upd-last-alreadydone}
			\State $\text{CAS}(L[x], l, (l_C.n, v, T))$
		\EndIf
	\EndFunction
	\Function{get-last}{x}
		\State $\fn{help}(x)$
		\State \Return $L[x]$
	\EndFunction
\end{algorithmic}

The use of binary search in line~\ref{help-binsch} warrants explanation. The predicate employed can obviously change its value in time. Thus, the binary search will return $v$ such that the predicate held at one point
for $v$ and didn't at another point for $v-1$. We will see that this is sufficient. Indeed, the result of the binary search will be important only in the cases when the value of the predicate doesn't change during
the search.

We will first prove two simple facts about \fn{update}:

\begin{lemma}
\label{lem-versions-increase}
For any two subsequent successful calls to hist.\fn{publish} with values $h_1$ and $h_2$, $h_1.v[i] \leq h_2.v[i]$ for both $i$.
\end{lemma}
\begin{proof}
Assume otherwise. Consider the first pair $p_1, p_2$ of consecutive calls to \fn{publish} that contradicts the Lemma (see Figure~\ref{fig-versions-increase}). The only place \fn{publish} is called is line \ref{upd-publish}
in \fn{update}.
$p_2$ had its most recent call to \fn{get-current} (denoted by $g_2$) in line~\ref{upd-getown} sometime earlier. No successful publish could have occured between $g_2$ and $p_2$; otherwise $p_2$
would have failed.
Thus, $p_1$ must have occured before $g_2$ and so its execution of lines \ref{upd-getch-0} and \ref{upd-getch-1} (denoted by ``read children'' on the diagram) must have occured strictly earlier than
the same for the second call. This together with monotonicity of versions in children delivers the contradiction.
\end{proof}

\begin{figure}
\begin{tikzpicture}
	\atomic{prev_getch}{read children}{1}{1}
	\atomic{prev_pub}{$p_1$}{3}{1}

	\atomic{get_ver}{$g_2$}{5}{0}
	\atomic{getch}{read children}{7}{0}
	\atomic{publish}{$p_2$}{9}{0}

	\draw[dashed] (0,-1) to (10,-1);
	\draw[dashed] (0,0) to (10,0);
	\draw (0.5,-1) to (3.5,-1);
	\draw (4.5,0) to (9.5,0);

	\draw[arrows={-triangle 45}] (prev_pub) to (get_ver);
\end{tikzpicture}
\caption{Diagram for proof of Lemma~\ref{lem-versions-increase} ($p_i$ -- call to \fn{publish}; $g_i$ -- call to \fn{get-current} in line~\ref{upd-getown} that corresponds to $p_i$)}
\label{fig-versions-increase}
\end{figure}

\begin{lemma}
	For every execution of \fn{write-and-f$_i$}: during execution of lines \ref{updcall1} and \ref{updcall2} at least one successful call to \fn{update} occurs.
\end{lemma}
\begin{figure}
\begin{tikzpicture}
	\atomic{our_gv1}{$g_1$}{5.5}{0}
	\atomic{bad_pub1}{$p^S_3$}{6.5}{1}
	\atomic{our_pub1}{$p^F_1$}{7.5}{0}
	\draw[arrows={-triangle 45}] (our_gv1) -- (bad_pub1);
	\draw[arrows={-triangle 45}] (bad_pub1) -- (our_pub1);
	\atomic{bad_gv2}{$g_4$}{7}{2}
	\draw[arrows={-triangle 45}] (bad_pub1) -- (bad_gv2);
	\atomic{our_gv2}{$g_2$}{9.5}{0}
	\atomic{bad_pub2}{$p^S_4$}{11}{2}
	\atomic{our_pub2}{$p^F_2$}{12}{0}
	\draw[arrows={-triangle 45}] (our_gv2) -- (bad_pub2);
	\draw[arrows={-triangle 45}] (bad_pub2) -- (our_pub2);

	\draw[dashed] (2.5,0) -- (13,0);
	\draw (3,0) -- (12.5,0);

	\draw[dashed] (5.5,-1) -- (7.5,-1);
	\draw (6,-1) -- (7,-1);

	\draw[dashed] (6,-2) -- (12,-2);
	\draw (6.5,-2) -- (11.5,-2);
\end{tikzpicture}
\caption{Consequences of two failed \fn{update} calls ($g_i$ -- call to \fn{get-current} in line~\ref{upd-getown}, $p^S_{i}$ -- successful call to \fn{publish} in line~\ref{upd-publish}, $p^F_{i}$ -- failed call to \fn{publish} in line~\ref{upd-publish})}
\label{fig-only-two-pubs}
\end{figure}

\begin{proof}
	If one of the calls to \fn{update} from the lines in question suceeded, then the Lemma trivially holds. Thus we assume that they have both failed. The situation is depicted
	in Figure~\ref{fig-only-two-pubs}. $g_1, p^F_1, g_2 \text{ and } p^F_2$ correspond to these two failed calls to \fn{update}. In order for $p^F_1$ to fail, a \fn{publish} must have
	succeeded between $g_1$ and $p^F_1$ -- let's call it $p^S_3$. Similarly, a \fn{publish} must have suceeded between $g_2$ and $p^F_2$. Alas, this \fn{publish} (denoted $p^S_4$) must have had
	its corresponding call to \fn{get-current} (denoted $g_4$) after $p^S_3$. The call to \fn{update} that called $g_4$ and $p^S_4$ satisfies the Lemma's conclusion.
\end{proof}

Let us consider an execution of \fn{write-and-f$_i$}. Let $n_c, v_c, T_c$ be the return value of the corresponding call to $C[side(i)].\fn{write-and-f$_{child\_id(i)}$}$. Then by the time the call to
\fn{write-and-f$_i$} finishes, the most recently published node value has $v[side(i)] \geq v_c$. We will consider the first call to \fn{publish} that publishes such a node value as the serialization
point of the \fn{write-and-f}. Obviously, many calls to \fn{write-and-f} can then serialize at the same instant in time. We will order them by taking first these with $side(i) = 0$ in the order in which
their corresponding calls to \fn{write-and-f} happened in $C[0]$ and then those with $side(i) = 1$ in the corresponding order. The choice of order on elements of $C$ is arbitrary, albeit is reflected
in \fn{help}.

Before we prove that the implementation is correct, we need the following two lemmas about properties of the $L$ array.

\begin{lemma}
	If \fn{help} is called after the serialization point of $n^{\text{th}}$ call to \fn{write-and-f$_i$}, then $L[i].n \geq n$ after the call to \fn{help} completes. \label{last-is-complete}
\end{lemma}
\begin{proof}
	We will prove this Lemma by induction on $n$. Let $p(v)$ be the time the successfull call to \fn{publish}$(v, \cdot)$ occured. Consider an \fn{update$_i$} that was serialized at $p(v)$.
	For every $u \equiv i (\mod N)$, a call to $\fn{help}(i)$ starts and finishes between $p(u-1)$ and $p(u)$. Thus, there is such a call that starts and finishes between $p(v)$ and $p(v+N)$.
	Without loss of generality we can assume that the call to \fn{help} from the Lemma's hypothesis executes in this time interval. In that case, the binary search from line~\ref{help-binsch} will not see any
	$\fn{get}(x)$ for $x \geq v-1$ returning \none and will	return version $v$. For the same reason, both \fn{get}s will succeed.

	We will now use the inductive hypothesis: at $p(v)$, $j-1^{\text{th}}$ call to \fn{write-and-f$_i$} has completed, so a call to $\fn{help}(i)$ has completed between the serialization point of that call to \fn{write-and-f$_i$}
	and $p(v)$. Thus $L[i].n \geq n-1$ from $p(v)$ on. If the condition in line~\ref{upd-last-alreadydone} is not satisfied or CAS in the next line fails, the Lemma hold.
	Otherwise, by the time CAS succeeds the Lemma will obviously hold.

\end{proof}

\begin{lemma}
	If a call to $\fn{help}(i)$ sets $L[i]$ to $n, v, T$, then by the time the call has finished: \label{last-is-correct}
\begin{enumerate}
	\item \fn{write-and-f$_i$} has been called at least $n$ times
	\item Let $n_c, v_c, T_c$ be the return value of the $n^{\text{th}}$ call to $C[side(i)].\fn{write-and-f}_{child\_id(p)}$. Let $v'$ and $h_new$ be the version and node value published at the serialization point of the $n^{\text{th}}$
		\fn{write-and-f$_i$} and $h_old$ be the node value published with version $v'-1$. Then $v = v'$ and:
		\begin{itemize}
			\item If $side(p) = 0$ then $T = f(T_c, h_{old}.T[1])$
			\item If $side(p) = 1$ then $T = f(h_{new}.T[0], T_c)$
		\end{itemize}
\end{enumerate}
\end{lemma}
\begin{proof}
	For the first part, we need to note that by the time line~\ref{get-last-recurse} executes, $n^{\text{th}}$ \fn{write-and-f$_{child\_id(i)}$} has already executed in the appropriate child.
	Let us note that if the binary search fails or returns version different from $v'$, \fn{help} will exit early. Indeed, if the binary search returns a different $v$, $\fn{get}(v-1)$ will fail.
	This immediately proves the first part of the Lemma and shows that $h_{old}$ and $h_{new}$ in the code have the same values as $h_{old}$ and $h_{new}$ in the Lemma, which proves the second part.
\end{proof}

\begin{theorem}
	The operations in multi-process write-and-f-array serialize to the single-process specification with serialization point of \fn{write-and-f} as posited earlier.
\end{theorem}
\begin{proof}
	Let us choose the call to \fn{get-current} as the serialization point of \fn{read}. 
	We will first prove that these are correct serialization points for \fn{write-and-f} and \fn{read}.
	Consider the $n^{\text{th}}$ call to \fn{write-and-f$_i$}. Let $n', v', T'$ be its return value. As this triple was read from $L[i]$, we can use Lemmas~\ref{last-is-complete} and~\ref{last-is-correct} to show that
	$n' = n$ (note that executions of \fn{write-and-f$_i$} for the same $i$ are disjoint in time). From Lemma~\ref{last-is-correct} we get that $v'$ is equal to the
	version published at serialization point of the call to \fn{write-and-f}. This obviously implies that the \fn{version} is nondecreasing for subsequent calls to \fn{write-and-f} and \fn{read}.
	As only the calls to \fn{write-and-f} that serialize on a single \fn{publish} return equal $v$, no \fn{read} call can serialize in between. This proves both properties required from \fn{version}. We still need to prove,
	that $T'$ and values returned by \fn{read} are consistent with posited serialization order.

	Consider a set of \fn{write-and-f} calls that are serialized together. Let $h_{new}$ and $h_{old}$ be the history elements published, respectively, at the serialization point and most recently before it. Let $a_i$ be the sequence of return values
	of $C[0].\fn{write-and-f}$ calls corresponding to \fn{write-and-f} calls in question, in serialization order. Let $b_i$ be a similar sequence for $C[1]$. By Lemma~\ref{lem-versions-increase} these are exactly the calls to
$C[i].\fn{write-and-f}$ serialized with versions in $\left(h_{old}.v[i], h_{new}.v[i]\right]$. By Lemma~\ref{last-is-correct} the sequence returned
by the \fn{write-and-f} calls is $f(a_0, b_0), $ $f(a_1, b_0),$ $\ldots, $ $f(h_{new}.T[0], b_1), $ $f(h_{new}.T[0], b_2), \ldots$ in the order of serialization. 
Additionally, $f(h_{new}.T[0], h_{new}.T[1])$ is the value that would be returned by any \fn{read} calls until the next serialization point of \fn{write-and-f}.
This suffices to show that the return values of consecutive calls to \fn{write-and-f} are actually the results of applying the updates.
This proves that values returned by \fn{get} and \fn{update} are correct wrt. posited serialization order.

	We still need to show that we can serialize \fn{get-last} calls correctly. Lemma~\ref{last-is-correct} implies that two calls to $\fn{get-last}(i)$ will never return different triples with equal $n$. As \fn{write-and-f} returns
	the result of a call to \fn{get-last}, we just need to show that \fn{get-last}$(i)$ can be serialized so that the $n$ in its return value (let's call it $n$) is the number of previously serialized calls to \fn{write-and-f$_i$}. From
	Lemma~\ref{last-is-complete} we get that $n$ is at least the number of calls to \fn{write-and-f$_i$} that serialized before \fn{get-last} started. From Lemma~\ref{last-is-correct} we know that $n$ was at most the number of such calls
	that serialized before \fn{get-last} finished. Thus, there is a point in the execution interval of \fn{get-last} when $n$ is equal to the number of such calls that have already serialized. We choose any such point in time to be the serialization
	point of the call to \fn{get-last}.
\end{proof}

The structure uses $O(N\log{}N)$ memory. A \fn{get-last} operation takes $O(\log^2 N)$ time, a \fn{read} operation takes $O(1)$ time and a \fn{write-and-f} operation takes $O(\log^3 N)$ time.

We can construct a fetch-and-add object for $P$ threads by using a write-and-f-array of size $P$ with addition as the operation $f$. Every thread is assigned an element in the array, and modifies only that element. This gives us
a fetch-and-add object for $P$ threads with $O(\log^3 P)$ step complexity and $O(P\log P)$ memory complexity. Additionally, the object implements a method that retrieves the
current value in $O(1)$ time.

\section{Implementation}
A fetch-and-add object was implemented using the mentioned construction from the write-and-f object. The implementation is in C++ and uses the C++11 standard library support for atomic operations.
The implementation modifies the object described in this paper so as to constrain the size of CAS object required to no more than twice the size of a version variable. It uses 64-bit versions and
thus requires 128-bit CAS objects. Thus, it runs on 64-bit Intel architecture processors, but not on 32-bit ones.

Unfortunately, the C++ standard library support for atomic operations doesn't allow atomic reads of a part of a larger atomic variable (specifically, reads of 64-bit halves of a 128-bit variable).
As 128-bit atomic reads on amd64 are very costly
(they use a 128-bit wide CAS), limiting their number was very important for the efficiency's sake. Thus, the implementation makes an unwarranted assumption that an atomic variable
has the same memory layout as a normal, non-atomic variable of the same size. This assumption holds for gcc VERSION on amd64.

The correctness of the implementation was tested experimentally both on real hardware and by using the Relacy Race Detector\ref{rrd}. Relacy Race Detector is a framework for testing multi-threaded programs in C++11.
It substitues its own
implementation of synchronization primitives and atomic variables and runs user-supplied tests multiple times, with various interleaving of threads.
It can detect data races on non-atomic variables, deadlock conditions, and failed user-supplied invariant and assertion
checks. One notable feature is support for reduced consistency operations -- Relacy can simulate acquire/release semantics, as specified in the C++11 atomic variables library.
Our fetch-and-add implementation can be compiled both to run on bare hardware and to run in Relacy. The use of Relacy not only privided confidence about correctness of the implementation, but also allowed us to downgrade
consistency guarantees of some writes.

The implementation was benchmarked on a machine with 4 12-core AMD Opteron 6174 processors.
The benchmark created a fetch-and-add object and started a preset number of threads. Each thread incremented the counter in a loop, counting its operations locally. After 20 seconds elapsed, all threads were signalled
to stop and their operation counters were summed. For comparison, the same benchmark was run with the fetch-and-add object implemented by a simple read-modify-write loop. All benchmarks were run when the machine was
minimally loaded (had 5 minute load average smaller than 0.5 at the start of the benchmark).

The reuslts of the benchmark are shown in Figure~x. The figure contains a plot of average time it takes to complete one operation in function of the number of concurrently executing threads.
The horizontal scale of the figure is proportional to the square of the logarithm of number of threads. The reason for this is that the optimistic time complexity
of our fetch-and-add object is $\Theta(\log^2 P)$. We can see that the graph for our implementation approximates a straight line up to XX threads, where it starts to diverge upward. This peculiar divergence disappears
if we assign one core to each thread and force it to run only there (by setting CPU affinity). Other affinity settings yielded a graph similar to one of these two. We were unable to explain this behaviour. It might
be worth noting that the anomaly in the no-affinity case happens when it starts using all 4 physical CPUs.

It can be easily seen from the graph that our fetch-and-add object is far from practical. It is about 100 times slower than the naive lock-free implementation for small numbers of threads and about 40 times slower
for larger numbers of threads. We believe that this implementation is suboptimal, although achieving similar performance as the naive implementation seems impossible. One change that could improve performance
of this implementation is changing the arity of the tree. The construction can easily be adapted to trees with larger arity and this would decrease the number of expensive CAS operations, while increasing the number
of atomic reads, which are comparatively cheap on Intel processors.

\section{Extensions and Open Problems}

Our implementation of write-and-f-array can be modified by splitting the work done by \fn{help} into $O(\log N)$ separate pieces (updates on consecutive levels of the tree) and running only a single such piece in \fn{update}.
If we then enlarge the history size to $O(N\log N)$, Lemma~\ref{last-is-complete} still holds. This modification increases the memory complexity to $O(N\log^2N)$ and decreases the step complexity of
\fn{write-and-f} to $O(\log^2P)$.

The structure can be straightforwardly modified to use LL/CS instead of CAS, albeit it then requires the ability to have two outstanding LLs at the same time. The modified version can be further modified to use
bounded version numbers, from a range of size $O(N)$.

We can also modify the history structure and the $L$ array so that we need CAS objects of size $O(\log M) + O(1)$, independently of the size of \T.

We pose two problems:

Our implementation of write-and-f-array is single-writer, so it's natural to consider the multi-writer write-and-f-array. Can we construct a multi-writer write-and-f-array in subquadratic memory with similar step complexities of the operations?

The memory consumption of write-and-f-array implemented with atomic registers and CAS or LL/CS objects is bounded from below by a linear function of the number of processes that can execute \fn{write-and-f} concurrently\cite{lowerbound}, so it must be $\Omega(N)$. Can we achieve lower memory and/or step complexities?

\section{Acknowledgements}
The author would like to thank dr. Grzegorz Herman for many fruitful discussions and help while preparing this paper. I would also like to thank Szymon Acedański for his help in improving the presentation of this paper.

This work was supported by the Polish Ministry of Science and Higher Education program ``Diamentowy Grant''.

\newpage

\begin{thebibliography}{9}
	\bibitem{snapshot-impl} Riany Y., Shavit N., Touitou D. ``Towards a practical snapshot algorithm'', Theoretical Computer Science 269, pp. 163--201 (2001)
	\bibitem{snapshot-idea} Afek Y., Attiya H., Dolev D., Gafni E., Merritt M., Shavit N. ``Atomic snapshots of shared memory'', Journal of the ACM 40, pp. 873--890 (1993)
	\bibitem{write-and-snap} Afek Y., Weisberger E. ``The instancy of snapshots and commuting objects'', Journal of Algorithms 30.1, pp. 68--105 (1999)
	\bibitem{closed-object} Chandra T. D., Jayanti P., Tan K. ``A polylog time wait-free construction for closed objects'', Proceedings the 17th PODC, pp. 287--296 (1998)
	\bibitem{f-array} Jayanti P. ``f-arrays: Implementation and applications'', Proceedings of the 21st PODC, pp. 270--279 (2002)
	\bibitem{ellen-fai} Ellen F., Ramachandran V., Woelfel P. ``Efficient Fetch-And-Increment'', Lecture Notes in Computer Science 7611, pp. 16--30 (2012)
	\bibitem{lowerbound} Fich F. E., Hendler D., Shavit N. ``Linear lower bounds on real-world implementations of concurrent objects'', Foundations of Computer Science, pp. 165--173 (2005)
	\bibitem{faa-sync} Freudenthal, E., Gottlieb, A. ``Process coordination with fetch-and-increment'', Proceedings of ASPLOS-IV, pp. 260–-268 (1991)
	\bibitem{faa-queue} Goodman, J., Vernon, M., Woest, P. ``Efficent synchronization primitives for large-scale cache-coherent multiprocessors'', Proceedings of ASPLOS-III, pp. 64-–75 (1989)
\end{thebibliography}

\end{document}